---
title: "Report"
author: "Anthony Ebert"
date: "22/08/2019"
output: pdf_document
bibliography: link.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
```

We are interested in finding the closest mechanistic model $q(\cdot|\lambda)$, parameterised by $\lambda$ to a known statistical model $p(\cdot|\theta)$, where $\theta$ is known. 

$$
\begin{aligned}
KL[q(\cdot)||p(\cdot|\theta)] &= \sum_{y \in \mathcal{Y}} q(y) \log \frac{ q(y) }{p(y|\theta)} \\
&= \log[z(\theta)] - \sum_{y \in \mathcal{Y}} q(y) \left[ \sum_i \theta_i s_i(y) \right] + \sum_{y \in \mathcal{Y}} q(y) \log q(y) 
\end{aligned}
$$

# The relative importance of the likelihood and entropy

```{r}
set.seed(1)

library(StartNetwork)

library(parallel)
cl <- makeCluster(detectCores())

param_range_large <- rep(seq(0.025, 0.975, by = 0.025), 20)

x1 <- parSapply(cl, param_range_large, er_KL, nl = 5, pl = 0.3, include_entropy = TRUE, replicates = 1000)
x2 <- parSapply(cl, param_range_large, er_graphs_KL, nl = 5, pl = 0.3, include_entropy = TRUE, replicates = 1000)
```


```{r, fig.cap="KL divergence calculation with entropy"}
library(ggplot2)


df <- data.frame(x = c(x1, x2), parameter = rep(param_range_large, 2), summary_statistic = c(rep("a", length(param_range_large)), rep("b", length(param_range_large))))

ggplot(df) + aes(x = parameter, y = x, group = parameter, col = summary_statistic) + geom_boxplot()

```

```{r, fig.cap = "KL divergence calculation without entropy"}


df <- data.frame(x = x2, parameter = param_range_large)

ggplot(df) + aes(x = parameter, y = x, group = parameter) + geom_boxplot()
```


```{r, fig.cap="KL divergence calculation with entropy"}


df <- data.frame(x = x3, parameter = param_range_small)

ggplot(df) + aes(x = parameter, y = x, group = parameter) + geom_boxplot()
```


```{r, fig.cap="KL divergence calculation without entropy"}

df <- data.frame(x = x4, parameter = param_range_small)

ggplot(df) + aes(x = parameter, y = x, group = parameter) + geom_boxplot()

```


# Attempt with preferential attachment model

The preferential attachment model has the following likelihood function:

$$
\begin{aligned}
P(k | \rho) &= \frac{(\rho - 1) \Gamma(k) \Gamma(\rho)}{\Gamma(k + \rho)}
\end{aligned}
$$

In this case there is no simple set of summary statistics. 

```{r}

x <- sapply(seq(2.5, 5.5, by = 0.1), ba_KL, pl = 4.5)

ggplot2::qplot(seq(2.5, 5.5, by = 0.1),x, xlab = "power parameter", ylab = "KL divergence") + ggplot2::geom_vline(aes(xintercept = x), data = data.frame(x = 4.5), col = "red")
```








